### 2021-02-19 architecture.tex
Respond to MS's notes about Architecture

The architecture is designed so that the recommender can be trained
on the results of \gls{atp} executions on various problems with random precedences.
Since the recommender contains a neural network,
it is parameterized by weight tensors
whose values can be trained by gradient descent.
\todo{Ensure we have defined ``gradient descent''.}
\todo{MS: chapu ucel (a vidim, ze dalsi veta navazuje na ``parameters'', ale je to kostrbate.
Recommender = system, chapu jako program. Programy vetsinou neparametrizujeme tenzorama a gradient descent zije na jine urovni,
nez aby parametrizovane programy primo ucil. Uceni je spis nejaka technologie, kterou recommender pouziva, aby ... ``produces precedences
that are likely to yield a successful proof''.  }
The goal of the training is to find parameter values
such that the recommender produces precedences
that are likely to yield a successful proof in few iterations of the saturation loop.
\todo{FB: Too vague? MS: v tomhle momente mi to prijde dobry!}

The recommender consists of modules that perform specific subtasks (see \cref{fig:architecture}):
\begin{enumerate}
\item Graphifier\todo{MS: Zatim nevidim Graphifier ve \cref{fig:architecture}.} converts the input \gls{cnf} problem into a problem graph.
\item \gls{gcn} converts a problem graph into symbol embeddings.
%The \gls{gcn} contains one or more convolution layers.
\item Projection converts the symbol embeddings into a vector of symbol costs.
\item Sorting converts the symbol costs into a symbol precedence.
It is only used in generating mode.
\item Loss function converts the symbol costs and a pair of precedences into a loss value.
It is only used in training mode.
\end{enumerate}
\todo{MS: tohle zatim chapu jako predbezny nastrel. Leda by uz pred tim bylo vysvetleno:
embedding, \gls{gcn}/convolution layers, ... }

@@

As the first step of the recommender processing pipeline,
the input problem is converted from a \gls{cnf} representation
to a \emph{heterogeneous directed graph}.\footnote{Data mining literature often uses the term \gls{hin} \cite{Shi2015} for heterogeneous graphs.
We prefer to conform to the terminology common in literature studying \glspl{gnn}.}
\todo{FB: Consider removing the footnote because it is basically useless.}
\todo{MS: Footnote asi remove, ale neexistuje citace pro ``heterogeneous directed graph''?}
Each of the nodes is labeled with a node type,
and each edge is labeled with an edge type,
informing\todo{MS: maybe better ``defining''?} the heterogeneous nature of the graph.
Each node corresponds to one of the elements that constitute the \gls{cnf} formula,
such as a clause, an atom or a predicate symbol.
Each such category of elements corresponds to one node type.
The edges represent the (oriented) relations between the elements,
for example the incidence relation between a clause and one of its (literal's) atoms,
or the relation between an atom and its predicate symbol.
\Cref{fig:CnfSchema} shows the types of nodes and edges used in our graph representation.
\Cref{fig:GcnExample} shows an example of a graph representation of a simple problem.

@@

The graph representation has the following properties:
\begin{itemize}
\item Lossless: The original problem can be faithfully reconstructed from the corresponding graph representation
(up to logical equivalence).
\item Signature agnostic: Renaming the symbols and variables in the input problem yields an isomorphic graph.
\item For each relation $r$, its inverse $\inv{r}$ is also present in the graph,
typically represented by a different edge type.
\todo{MS: tohle jsem nepochopil. Nemuzeme rict, ze (az na nejakou vyjimkou) jsou edges v obou smerech?}
\item A singleton node of type \ntype{problem} is connected to all the clauses of the problem.
\item The polarity of the literals is expressed by the type of the edge (\epos{} or \eneg{})
connecting the respective atom to the clause it occurs in.
\item For every non-equality atom and term, the order of its arguments is captured by a sequence of \ntype{argument} nodes chained by edges \cite{Rawson2020}.
\item The two operands of equality are not ordered.
This reflects the symmetry of equality.
\item Term-sharing is employed:\todo{MS: Mozna presneji ``perfect sub-expression sharing''? 
Predpokladam ze dva ruzne ground literaly p(c) tam budou taky jen jendnou?}
Redundant atoms and terms share a node representation \cite{}.
Note that since each variable is bound by a clause,
ground terms are shared across clauses,
but non-ground terms are only shared within the context of a clause.
\todo{FB: Remove as of little interest?}
\end{itemize}
\todo{FB: Reference appendix if we describe the representation in more detail there.}
\todo[inline]{MS: Filosoficka k ``can be represented''; mozna by se mohlo nekde objevit:
Snazime se do struktury HIN ``otisknout'' celou strukturu CNF (jak jsem rikal,
znamena to, ze zobrazeni je proste), abychom siti nevzali moznost ``vedet o problemu
uplne vsechno'' (az na veci, na kterych nezalezi jako poradi literaly v klauzulich, etc).
Ta filosoficka poznamka je o tom, ze pro dobre uceni tohle vlastne vubec nemusi byt treba,
ze nejaky ``hrubsi otisk'' by mohl stacit tez (a byt treba mensi), ale my nevime, jak takovy zvolit. }

@@

For each symbol in the problem,
we seek to find a vector representation, i.e., an \emph{embedding},
that captures the symbol's properties that are relevant
for correctly ranking the symbol in the observed symbol precedences.
The symbol embeddings are modeled\todo{MS: embeddings are ``output''? } by a \gls{rgcn} \cite{Schlichtkrull2017},
\todo{Can I use the term \gls{rgcn} when I use a modified version? Namely, the activation is called earlier. MS: sounds good.}
which is a stack of graph convolutional layers.
Each layer consists of a collection of differentiable modules---one module per edge type.
The computation of the \gls{gcn} starts with assigning each node an initial embedding
and then repeatedly updates the embeddings by passing them through convolutional layers.
\todo{Ensure terms are clear (preferably defined): embedding, vector, trainable vector, feature vector.}

@@

One pass through the convolutional layer
updates the node embeddings by passing a message along each of the edges.
\todo{MS: ``message'' occurs for the first time. 
Maybe ``a message'', or explain beforehand.}
For an edge of type $r \in \mathcal{R}$ going from source node $s$ to destination node $d$ at layer $l$,
the message is composed by converting the embedding of the source node $h_s^{(l)}$
using the module associated with the edge type $r$.
In the simple case that the module is a dense layer with weights $W_r^{(l)}$,
\todo{MS: hnidopich nevi, co to znamena ``dense layer''}
the message is $h_s^{(l)} W_r^{(l)}$.
Each message is then divided by a normalization constant
$c_{s,d} = \sqrt{\card{\mathcal{N}_s^r}} \sqrt{\card{\mathcal{N}_d^r}}$ \cite{kipf2017semisupervised},
where $\mathcal{N}_n^r$ is the set of neighbors of node $n$ under the relation $r$.
\todo{FB: Mention that we mean both in- and out-neighbors?}
Once all the messages are computed,
they are aggregated by the destination nodes to form the new node embedding.
Each node $d$ aggregates all the incoming messages of a given edge type $r$ by summation,
then passes the sum through an activation function $\sigma$ such as the \gls{relu},
and finally aggregates the messages across the edge types by summation,
yielding the new embedding $h_d^{(l+1)}$.
The following formula captures the update of embedding of node $d$:
$$
h_d^{(l+1)} =
\sum_{r \in \mathcal{R}} \sigma \Parentheses{\sum_{j \in \mathcal{N}_s^r} \frac{1}{c_{s,d}} h_s^{(l)} W_r^{(l)}}
$$
%$$
%h_i^{(l+1)} =
%\mathrm{LayerNorm} \Parentheses{h_i^{(l)} + \sum_{r \in \mathcal{R}} \sigma \Parentheses{\sum_{j \in %\mathcal{N}_i^r} \frac{1}{c_{ji}} h_j^{(l)} W_r^{(l)}}}
%$$
% Inspiration: https://ufal.mff.cuni.cz/~straka/courses/npfl114/1920/slides.pdf/npfl114-07.pdf - slide 27 - Transformer
%\todo{FB: Mention inspiration: Transformer. See Straka's slides.}

>>>>

The architecture is designed so that the recommender can be trained
on the results of \gls{atp} executions on various problems with random precedences.
Since the recommender contains a neural network,
it is parameterized by weight tensors
whose values can be trained by gradient descent.
\todo{Ensure we have defined ``gradient descent''.}
\todo{MS: chapu ucel (a vidim, ze dalsi veta navazuje na ``parameters'', ale je to kostrbate.
Recommender = system, chapu jako program. Programy vetsinou neparametrizujeme tenzorama a gradient descent zije na jine urovni,
nez aby parametrizovane programy primo ucil. Uceni je spis nejaka technologie, kterou recommender pouziva, aby ... ``produces precedences
that are likely to yield a successful proof''.  }
The goal of the training is to find parameter values
such that the recommender produces precedences
that are likely to yield a successful proof in few iterations of the saturation loop.

The recommender consists of modules that perform specific subtasks,
each of which is described in detail in one of the following sections (see also \cref{fig:architecture}):
\begin{enumerate}
\item Graphifier converts the input \gls{cnf} problem into a problem graph.
\item \gls{gcn} converts a problem graph into symbol embeddings.
%The \gls{gcn} contains one or more convolution layers.
\item Projection converts the symbol embeddings into a vector of symbol costs.
\item Sorting converts the symbol costs into a symbol precedence.
It is only used in generating mode.
\item Loss function converts the symbol costs and a pair of precedences into a loss value.
It is only used in training mode.
\end{enumerate}
\todo{MS: tohle zatim chapu jako predbezny nastrel. Leda by uz pred tim bylo vysvetleno:
embedding, \gls{gcn}/convolution layers, ... \\
FB: Staci predestrit, ze detaily vysvetlim pozdeji?}

@@

As the first step of the recommender processing pipeline,
the input problem is converted from a \gls{cnf} representation
to a \emph{heterogeneous (directed) graph} \cite{Zhou2018}.
% Zhou uses the term "heterogeneous graph".
%\footnote{Data mining literature often uses the term \gls{hin} \cite{Shi2015} for heterogeneous graphs.
%We prefer to conform to the terminology common in literature studying \glspl{gnn}.}
Each of the nodes is labeled with a node type,
and each edge is labeled with an edge type,
defining the heterogeneous nature of the graph.
Each node corresponds to one of the elements that constitute the \gls{cnf} formula,
such as a clause, an atom or a predicate symbol.
Each such category of elements corresponds to one node type.
The edges represent the (oriented) relations between the elements,
for example the incidence relation between a clause and one of its (literals') atoms,
or the relation between an atom and its predicate symbol.
\Cref{fig:CnfSchema} shows the types of nodes and edges used in our graph representation.
\Cref{fig:GcnExample} shows an example of a graph representation of a simple problem.

@@

The graph representation has the following properties:
\begin{itemize}
\item Lossless: The original problem can be faithfully reconstructed from the corresponding graph representation
(up to logical equivalence).
\item Signature agnostic: Renaming the symbols and variables in the input problem yields an isomorphic graph.
\item For each relation $r$, its inverse $\inv{r}$ is also present in the graph,
typically represented by a different edge type.
\todo{MS: tohle jsem nepochopil. Nemuzeme rict, ze (az na nejakou vyjimkou) jsou edges v obou smerech?\\
FB: Chci vyjasnit, ze inverzni edge ma jiny edge type.}
\item A singleton node of type \ntype{problem} is connected to all the clauses of the problem.
\item The polarity of the literals is expressed by the type of the edge (\epos{} or \eneg{})
connecting the respective atom to the clause it occurs in.
\item For every non-equality atom and term, the order of its arguments is captured by a sequence of \ntype{argument} nodes chained by edges \cite{Rawson2020}.
\item The two operands of equality are not ordered.
This reflects the symmetry of equality.
\item Perfect sub-expression sharing:
\todo{MS: Predpokladam ze dva ruzne ground literaly p(c) tam budou taky jen jendnou?\\
FB: Literaly nemaji nody. Atomy maji nody a polarita literalu je zakodovana v typu hrany, ktery jej spojuje s klauzuli.}
Redundant atoms and terms share a node representation \cite{}.
Note that since each variable is bound by a clause,
ground terms are shared across clauses,
but non-ground terms are only shared within the context of a clause.
\todo{FB: Remove as of little interest?}
\end{itemize}
\todo{FB: Reference appendix if we describe the representation in more detail there.}
\todo[inline]{MS: Filosoficka k ``can be represented''; mozna by se mohlo nekde objevit:
Snazime se do struktury HIN ``otisknout'' celou strukturu CNF (jak jsem rikal,
znamena to, ze zobrazeni je proste), abychom siti nevzali moznost ``vedet o problemu
uplne vsechno'' (az na veci, na kterych nezalezi jako poradi literaly v klauzulich, etc).
Ta filosoficka poznamka je o tom, ze pro dobre uceni tohle vlastne vubec nemusi byt treba,
ze nejaky ``hrubsi otisk'' by mohl stacit tez (a byt treba mensi), ale my nevime, jak takovy zvolit.\\
FB: Zkusim to popsat, kdyz zbyde cas.}
\todo{FB: Mention other encodings that have been proposed by Mirek and Michael.}

@@

For each symbol in the problem,
we seek to find a vector representation, i.e., an \emph{embedding},
that captures the symbol's properties that are relevant
for correctly ranking the symbol in the observed symbol precedences.
\todo{FB: What does "correct" mean?}
The symbol embeddings are output by a \gls{rgcn} \cite{Schlichtkrull2017},
which is a stack of graph convolutional layers.
Each layer consists of a collection of differentiable modules---one module per edge type.
The computation of the \gls{gcn} starts with assigning each node an initial embedding
and then repeatedly updates the embeddings by passing them through convolutional layers.
\todo{FB: Ensure terms are clear (preferably defined): embedding, vector, trainable vector, feature vector.}

@@

One pass through the convolutional layer
updates the node embeddings by passing a message along each of the edges.
For an edge of type $r \in \mathcal{R}$ going from source node $s$ to destination node $d$ at layer $l$,
the message is composed by converting the embedding of the source node $h_s^{(l)}$
using the module associated with the edge type $r$.
In the simple case that the module is a dense layer with weights $W_r^{(l)}$,
\todo{MS: hnidopich nevi, co to znamena ``dense layer''}
the message is $h_s^{(l)} W_r^{(l)}$.
Each message is then divided by a normalization constant
$c_{s,d} = \sqrt{\card{\mathcal{N}_s^r}} \sqrt{\card{\mathcal{N}_d^r}}$ \cite{kipf2017semisupervised},
where $\mathcal{N}_n^r$ is the set of neighbors of node $n$ under the relation $r$.
\todo{FB: Mention that we mean both in- and out-neighbors?}
Once all the messages are computed,
they are aggregated by the destination nodes to form the new node embedding.
Each node $d$ aggregates all the incoming messages of a given edge type $r$ by summation,
then passes the sum through an activation function $\sigma$ such as the \gls{relu},
and finally aggregates the messages across the edge types by summation,
yielding the new embedding $h_d^{(l+1)}$.
The following formula captures the update of embedding of node $d$:
$$
h_d^{(l+1)} =
\sum_{r \in \mathcal{R}} \sigma \Parentheses{\sum_{j \in \mathcal{N}_s^r} \frac{1}{c_{s,d}} h_s^{(l)} W_r^{(l)}}
$$
%$$
%h_i^{(l+1)} =
%\mathrm{LayerNorm} \Parentheses{h_i^{(l)} + \sum_{r \in \mathcal{R}} \sigma \Parentheses{\sum_{j \in %\mathcal{N}_i^r} \frac{1}{c_{ji}} h_j^{(l)} W_r^{(l)}}}
%$$
% Inspiration: https://ufal.mff.cuni.cz/~straka/courses/npfl114/1920/slides.pdf/npfl114-07.pdf - slide 27 - Transformer
%\todo{FB: Mention inspiration: Transformer. See Straka's slides.}
